---
title: "Weather and European Electricity Load: Econometric Analysis"
author: "Giang Tran"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 12,
  fig.height = 5,
  dpi = 120
)

# Paths relative to final-project/ (set working directory when knitting)
DATA_DIR <- "data"
FIG_DIR <- "report/figures"
FOCUS_COUNTRIES <- c("DE", "ES", "FI", "FR", "PL")
ANALYSIS_END <- as.Date("2024-12-31")
dir.create(FIG_DIR, showWarnings = FALSE, recursive = TRUE)
```

# Introduction

This notebook implements the econometric analysis for the final project: **the dynamic relationship between weather (temperature) and European electricity load**.

- **Data:** ERA5 reanalysis weather merged with ENTSO-E hourly load (`final-project/data/*.parquet`).
- **Main model:** ARDL (Autoregressive Distributed Lag).
- **Robustness:** Seasonal ARIMA with exogenous weather regressors (SARIMAX-style).
- **Focus countries:** `r paste(FOCUS_COUNTRIES, collapse = ", ")` (in-depth analysis).
- **Appendix:** Summary metrics for all countries in the cleaned panel.

## Data provenance

| Step | Script | Output |
|--------------------|--------------------------|--------------------------|
| Weather | `codes/03_download_weather.py` | ERA5 NetCDF in `era5_eu_data/` |
| Load | `codes/02_download_load.py` | `energy_country_level/hourly_load_country_*.parquet` |
| Merge | `codes/04_merge_data.py` | Integrated parquet files in `data/` |

Three analysis frequencies are available:

- `final_integrated_u_curve_dataset.parquet` - hourly
- `12-hour_integrated_u_curve_dataset.parquet` - 12-hour
- `daily_integrated_u_curve_dataset.parquet` - daily (main ARDL input)

The merge step maps weather to country centroids defined in `codes/04_merge_data.py` (21 EU countries).

```{r libraries, echo=FALSE}
required_pkgs <- c(
  "tidyverse", "arrow", "zoo", "xts", "dynlm", "lmtest", "sandwich",
  "forecast", "tseries", "urca", "fUnitRoots", "mgcv", "patchwork", "car",
  "gratia"
)
missing_pkgs <- required_pkgs[!vapply(required_pkgs, requireNamespace, quietly = TRUE, FUN. = logical(1))]
if (length(missing_pkgs) > 0) {
  stop(
    "Install missing packages: ", paste(missing_pkgs, collapse = ", "),
    call. = FALSE
  )
}

library(tidyverse)
library(arrow)
library(zoo)
library(xts)
library(dynlm)
library(lmtest)
library(sandwich)
library(forecast)
library(tseries)
library(urca)
library(fUnitRoots)
library(mgcv)
library(patchwork)
library(car)
library(gratia)
library(forcats)

options(scipen = 999)

source("R/analysis-helpers.R")
```

# Load and Clean Data

> **Purpose.** Read the daily parquet (used by EDA / ADF / OLS / ARDL) and the hourly parquet filtered to Germany only (used by section 6 SARIMAX); build derived columns (`log_load`, `temp_c2`, `is_weekend`, `month`) and export the cleaned daily CSV for submission.
>
> **Expected result.** ~38k daily rows, 21 countries, no `NA` in `log_load` / `temperature_c`, date range 2020-01-01 to 2024-12-31.
>
> **If not as expected.** Re-check parquet paths, missing centroid mapping in `codes/04_merge_data.py`, or `NA` on `load_mw` (rerun ENTSO-E download with the `ENTSOE_API_KEY` env var).
>
> **What to look at.** `nrow(daily_clean)`, country count, summary table of mean load and mean temperature per country, presence of `data/cleaned_daily_weather_load.csv`.

```{r load-data, echo=FALSE}
daily_raw <- read_parquet(file.path(DATA_DIR, "daily_integrated_u_curve_dataset.parquet"))

pop_annual <- read_csv(
  file.path(DATA_DIR, "eurostat_population_annual.csv"),
  show_col_types = FALSE
)

# Only DE hourly observations are needed for the hourly ARIMA robustness section.
hourly_raw <- arrow::open_dataset(
  file.path(DATA_DIR, "final_integrated_u_curve_dataset.parquet")
) %>%
  filter(country_code == "DE") %>%
  collect()

AVAILABLE_COUNTRIES <- sort(unique(daily_raw$country_code))

daily_clean <- daily_raw %>%
  mutate(
    date = as.Date(timestamp),
    country_code = as.character(country_code),
    year = as.integer(format(as.Date(timestamp), "%Y"))
  ) %>%
  left_join(pop_annual, by = c("country_code", "year")) %>%
  filter(
    date <= ANALYSIS_END,
    country_code %in% AVAILABLE_COUNTRIES
  ) %>%
  mutate(
    log_load = log(load_mw),
    temp_c2 = temperature_c^2,
    month = as.factor(format(timestamp, "%m")),
    weekday = weekdays(as.Date(timestamp)),
    is_weekend = as.factor(ifelse(weekday %in% c("Saturday", "Sunday"), 1, 0))
  ) %>%
  filter(
    is.finite(log_load),
    is.finite(temperature_c),
    is.finite(temp_c2)
  ) %>%
  arrange(country_code, date)

TARGET_LOAD_COL <- "load_mw"

# Submission dataset: cleaned daily panel for focus + all countries
cleaned_export <- daily_clean %>%
  select(
    country_code, date, year, population, load_mw, log_load, temperature_c, temp_c2,
    dewpoint_c, wind_speed_ms, solar_radiation_wm2, precipitation_mm, tcc,
    weekday, is_weekend, month
  )

write_csv(
  cleaned_export,
  file.path(DATA_DIR, "cleaned_daily_weather_load.csv")
)

cat("Countries in panel:", length(AVAILABLE_COUNTRIES), "\n")
cat("Daily observations:", nrow(daily_clean), "\n")
cat("Date range:", as.character(min(daily_clean$date)), "to", as.character(max(daily_clean$date)), "\n")
knitr::kable(
  daily_clean %>%
    group_by(country_code) %>%
    summarise(
      n_days = n(),
      mean_load_mw = round(mean(load_mw), 0),
      mean_temp_c = round(mean(temperature_c), 1),
      .groups = "drop"
    ),
  caption = "Daily panel summary by country"
)
```

```{r panel-summary-export, echo=FALSE}
panel_summary <- build_panel_country_summary(daily_clean, pop_annual)
write_csv(panel_summary, file.path(DATA_DIR, "panel_country_summary.csv"))
```

# Exploratory Data Analysis

> **Purpose.** Visually check load level / seasonality and the temperature-load U-curve before any formal test.
>
> **Expected result.** Clear annual cycle in load. U-shape in the temperature-load scatter for `DE`, `FR`, `PL` (both heating and cooling). Monotonic decreasing for `FI` (heating only). Cooling-dominant (rising) for `ES`.
>
> **If not as expected.** Check unit conversions in the weather merge (Kelvin vs Celsius in `codes/04_merge_data.py`) and country-code consistency.
>
> **What to look at.** Time-series facet plot, the country-by-country quadratic fit on the temperature-load scatter, and the `cor(temperature_c, load_mw)` table.

```{r eda-load-series, fig.height=6, echo=FALSE}
focus_daily <- daily_clean %>% filter(country_code %in% FOCUS_COUNTRIES)

ggplot(focus_daily, aes(x = date, y = load_mw, color = country_code)) +
  geom_line(linewidth = 0.3, alpha = 0.85) +
  facet_wrap(~country_code, ncol = 1, scales = "free_y") +
  labs(
    title = "Daily electricity load (focus countries)",
    x = NULL, y = "Load (MW)", color = "Country"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r eda-temp-load, fig.height=6, echo=FALSE}
ggplot(focus_daily, aes(x = temperature_c, y = load_mw, color = country_code)) +
  geom_point(alpha = 0.12, size = 0.4) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2, raw = TRUE), se = TRUE) +
  facet_wrap(~country_code, scales = "free_y") +
  labs(
    title = "Temperature vs load (quadratic fit)",
    x = "Temperature (°C)", y = "Load (MW)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

# Stationarity Tests (ADF, KPSS, PP)

> **Purpose.** Decide whether `log_load` and `temperature_c` are I(0) or I(1) using three complementary unit-root tests, then choose whether to model in levels or first differences.
>
> **Expected result.** **Levels:** ADF / PP fail to reject the unit-root null, KPSS rejects the stationarity null - verdict non-stationary. **First differences:** ADF / PP reject the unit-root null, KPSS does **not** reject the stationarity null - verdict stationary, series is declared I(1) and modelled in differences in section 5.
>
> **If not as expected.** If a single test disagrees, prefer KPSS as tie-breaker (its null is stationarity, so a non-rejection is strong evidence of stationarity). If still ambiguous, try a longer ADF augmentation order; if differences remain non-stationary, take second differences.
>
> **What to look at.** The combined table per series with columns `adf_p`, `pp_p`, `kpss_stat` vs `kpss_5pct_crit`, and the consolidated verdict `stationary_all_three`.

```{r stationarity-focus, echo=FALSE, fig.show='hide'}
adf_results <- map_dfr(FOCUS_COUNTRIES, function(cc) {
  sub <- daily_clean %>% filter(country_code == cc) %>% arrange(date)
  bind_rows(
    bind_cols(tibble(country = cc), stationarity_block(
      sub$log_load, "log_load_levels", adf_type = "c", kpss_type = "mu", pp_model = "constant"
    )),
    bind_cols(tibble(country = cc), stationarity_block(
      diff(sub$log_load), "d_log_load", adf_type = "nc", kpss_type = "mu", pp_model = "constant"
    )),
    bind_cols(tibble(country = cc), stationarity_block(
      sub$temperature_c, "temperature_c_levels", adf_type = "c", kpss_type = "mu", pp_model = "constant"
    )),
    bind_cols(tibble(country = cc), stationarity_block(
      diff(sub$temperature_c), "d_temperature_c", adf_type = "nc", kpss_type = "mu", pp_model = "constant"
    ))
  )
})

adf_results %>%
  select(country, series, augmentations, adf_stat, adf_p, pp_stat, pp_p, kpss_stat, stationary_all_three) %>%
  knitr::kable(
    digits = 4,
    caption = "ADF / PP / KPSS - focus countries (verdict requires all three to agree on stationarity)"
  )

write_csv(
  adf_results %>%
    select(country, series, augmentations, adf_stat, adf_p, pp_stat, pp_p,
           kpss_stat, stationary_all_three),
  file.path(DATA_DIR, "stationarity_focus_countries.csv")
)
```

**Interpretation.** Daily `log(load)` and `temperature_c` are typically non-stationary in **levels** (ADF / PP do not reject, KPSS rejects). After first-differencing both series become stationary (all three tests agree). This justifies the ARDL specification of section 5 in **first differences**.

# Baseline Static Models (OLS and GAM)

> **Purpose.** Fit a quadratic OLS U-curve and a GAM smooth, then run a full diagnostic suite (BG, BP, RESET, JB, VIF) to show why a static model is insufficient.
>
> **Expected result.** Negative `temperature_c` coefficient and positive `I(temperature_c^2)` coefficient (open-up U-curve). `R^2` typically above 0.6. Breusch-Godfrey `p < 0.01` at orders 1 **and** 7 (strong daily and weekly autocorrelation). Breusch-Pagan typically rejects (heteroskedastic). Ramsey RESET typically rejects (functional form misspecified). JB usually rejects (non-normal residuals). VIF for `temperature_c` and `I(temperature_c^2)` is high by construction.
>
> **If not as expected.** If `bg_p_order1 > 0.10` at **all** orders, the static OLS is already adequate. If the U-curve coefficients have the wrong sign, double-check the country filter and the weather merge.
>
> **What to look at.** Coefficient table; U-curve turning point `-beta1 / (2 * beta2)`; residual ACF; diagnostics table (BG, BP, RESET, JB, VIF).

```{r baseline-focus, echo=FALSE}
baseline_results <- map_dfr(FOCUS_COUNTRIES, function(cc) {
  sub <- daily_clean %>% filter(country_code == cc)
  m_ols <- fit_baseline_ols(sub)
  m_gam <- fit_baseline_gam(sub)

  tibble(
    country = cc,
    ols_r2 = summary(m_ols)$r.squared,
    ols_beta_temp = coef(m_ols)["temperature_c"],
    ols_beta_temp2 = coef(m_ols)["I(temperature_c^2)"],
    turning_point_c = u_curve_turning_point(m_ols),
    gam_r2 = summary(m_gam)$r.sq,
    ols_aic = AIC(m_ols),
    gam_aic = AIC(m_gam)
  )
})

baseline_results %>%
  knitr::kable(digits = 4, caption = "Baseline OLS/GAM - focus countries")

write_csv(baseline_results, file.path(DATA_DIR, "baseline_ols_focus.csv"))
```

```{r baseline-diagnostics-table, echo=FALSE}
baseline_diag_table <- map_dfr(FOCUS_COUNTRIES, function(cc) {
  sub <- daily_clean %>% filter(country_code == cc)
  m <- fit_baseline_ols(sub)
  diag <- baseline_diagnostics(m)
  bind_cols(tibble(country = cc), as_tibble(as.list(diag)))
})

baseline_diag_table %>%
  knitr::kable(
    digits = 4,
    caption = "Baseline OLS diagnostics - Breusch-Godfrey (BG), Breusch-Pagan (BP), Ramsey RESET, Jarque-Bera (JB), VIF on temperature_c"
  )

write_csv(baseline_diag_table, file.path(DATA_DIR, "baseline_diagnostics_focus.csv"))
```

```{r baseline-residual-acf, fig.height=8, echo=FALSE}
for (cc in FOCUS_COUNTRIES) {
  sub <- daily_clean %>% filter(country_code == cc)
  m <- fit_baseline_ols(sub)
  res <- residuals(m)
  max_lag <- min(60, length(res) - 1)
  acf_vals <- acf(res, lag.max = max_lag, plot = FALSE)$acf[, , 1]
  ci <- 1.96 / sqrt(length(res))

  p_acf <- ggplot(
    data.frame(lag = 0:max_lag, acf = acf_vals),
    aes(lag, acf)
  ) +
    geom_segment(aes(xend = lag, yend = 0), linewidth = 0.6) +
    geom_hline(yintercept = c(-ci, ci), linetype = "dotted", color = "red") +
    labs(title = paste("OLS residual ACF -", cc), x = "Lag (days)", y = "ACF") +
    theme_minimal()

  print(p_acf)
}
```

# Variable Selection: Country-by-Country OLS Specification

This section performs general-to-specific variable selection for each focus country on the **static OLS** in levels. The goal is to identify which weather regressors belong in each country's parsimonious model before moving to the dynamic ARDL specification in Section 6.

## Germany (DE)

```{r de-subset, echo=FALSE}
target_country <- "DE"

sub_df <- daily_clean %>%
  filter(country_code == target_country) %>%
  dplyr::select(
    all_of(TARGET_LOAD_COL), temperature_c, dewpoint_c, wind_speed_ms,
    solar_radiation_wm2, tcc, precipitation_mm, month, is_weekend
  ) %>%
  na.omit()
```

### Model 1: Full General Uncentered OLS Baseline

```{r de-model1-baseline, echo=FALSE}
protected_vars <- c("temperature_c", "I(temperature_c^2)", "month", "is_weekend")
candidate_vars_full <- c("dewpoint_c", "wind_speed_ms", "solar_radiation_wm2", "tcc", "precipitation_mm")
current_vars_m1 <- c(protected_vars, candidate_vars_full)
formula_m1 <- paste(TARGET_LOAD_COL, "~", paste(current_vars_m1, collapse = " + "))
model_ols_general <- lm(as.formula(formula_m1), data = sub_df)
vif(model_ols_general)
summary(model_ols_general)
```

#### Omit `dewpoint_c`, `solar_radiation_wm2` due to high collinearity

```{r de-211a}
linearHypothesis(model_ols_general, c("dewpoint_c = 0", "solar_radiation_wm2 = 0"))
```

```{r de-211b}
model_ols_reduced1 <- update(model_ols_general, . ~ . - dewpoint_c - solar_radiation_wm2)
summary(model_ols_reduced1)
```

#### Jointly significant test of all insignificant variables

```{r de-212}
linearHypothesis(model_ols_general, c("dewpoint_c = 0", "solar_radiation_wm2 = 0", "month02 = 0", "month11 = 0", "precipitation_mm = 0"))
```

#### Omit `month11` due to highest insignificant

```{r de-213}
linearHypothesis(model_ols_general, c("dewpoint_c = 0", "solar_radiation_wm2 = 0", "month11 = 0"))
sub_df_reduced <- sub_df %>%
  mutate(
    month_reduced = fct_collapse(as.factor(month), "01" = c("11"))
  )

model_ols_reduced2 <- update(
  model_ols_general, 
  . ~ . - dewpoint_c - solar_radiation_wm2 - month + month_reduced,
  data = sub_df_reduced
)

summary(model_ols_reduced2)
```

#### Omit `precipitation_mm` due to highest insignificant

```{r de-214}
linearHypothesis(model_ols_general, c("dewpoint_c = 0", "solar_radiation_wm2 = 0", "month11 = 0", "precipitation_mm = 0"))
model_ols_reduced3 <- update(
  model_ols_general, 
  . ~ . - dewpoint_c - solar_radiation_wm2 - month + month_reduced - precipitation_mm,
  data = sub_df_reduced
)

summary(model_ols_reduced3)
```

**Final model for Germany:** `load_mw ~ temperature_c + I(temperature_c^2) + is_weekend + wind_speed_ms + tcc + month_reduced` (November merged into January baseline).

### OLS Diagnostic Summary Table

```{r de-ols-summary, echo=FALSE}
coefs_m2 <- coef(model_ols_reduced2)
r_sq_m2 <- summary(model_ols_reduced2)$r.squared
sub_df_reduced$residuals_ols_reduced <- residuals(model_ols_reduced2)

summary_list_m2 <- data.frame(
  Country = target_country,
  Beta_Intercept = coefs_m2[1],
  Beta_T = coefs_m2[2],
  Beta_T2 = coefs_m2[3],
  R_Squared = r_sq_m2,
  row.names = NULL
)
knitr::kable(summary_list_m2, caption = "Reduced OLS Framework Summary Parameters")
```

### Model 2: Centered Specific OLS Framework

```{r de-ols-centered, echo=FALSE}
sub_df_reduced$temperature_c_centered <- scale(sub_df_reduced$temperature_c, scale = FALSE)
model_ols_centered <- update(
  model_ols_reduced3, 
  . ~ . - temperature_c - I(temperature_c^2) + temperature_c_centered + I(temperature_c_centered^2),
  data = sub_df_reduced
)
cat("Final specification VIF\n")
vif(model_ols_centered)
summary(model_ols_centered)
```

Key interpretations:

- The **Intercept** (57,211 MW) is the expected baseline load on a weekday in January at average temperature, with zero for other weather variables.
- **Weekend drop** (`is_weekend1`): load drops by ~9,776 MW on weekends - the classic signature of commercial/industrial shutdown.
- **Seasonality** (`month`): base load is generally lower in spring/summer. December is also lower (-2,908 MW) due to holiday-period demand reduction.
- **Cloud Cover** (`tcc`): each unit increase raises load by ~1,415 MW (increased lighting demand, reduced rooftop solar output).
- **Wind Speed** (`wind_speed_ms`): each m/s raises load by ~412 MW - likely the wind chill effect increasing heating demand.

### AIC Model Comparison

```{r de-AIC, echo=FALSE}
aic_output <- AIC(model_ols_centered, model_ols_reduced3)
knitr::kable(aic_output, caption = "Model Fitness Selection Matrix (Lower AIC wins)")
```

### GAM nonlinearity check (GTS-final controls)

```{r de-gam-gts, echo=FALSE}
model_gam_de <- gam(
  load_mw ~ s(temperature_c_centered, bs = "cr", k = 12) +
    is_weekend + month_reduced + wind_speed_ms + tcc,
  data = sub_df_reduced,
  method = "REML"
)

gam_aic_comparison_de <- tibble(
  model = c("Quadratic OLS (final, centred)", "GAM (spline on T, same controls)"),
  aic = c(AIC(model_ols_centered), AIC(model_gam_de)),
  r_squared = c(
    summary(model_ols_centered)$r.squared,
    summary(model_gam_de)$r.sq
  )
)

gam_aic_comparison_de %>%
  knitr::kable(digits = 4, caption = "AIC comparison: quadratic OLS vs GAM (Germany, static GTS stage)")

write_csv(gam_aic_comparison_de, file.path(DATA_DIR, "gam_aic_comparison_de.csv"))
```

```{r de-gts-export, echo=FALSE}
gts_ols_reduction_de <- build_gts_ols_reduction_de(
  model_ols_general, model_ols_reduced2, model_ols_reduced3
)
write_csv(gts_ols_reduction_de, file.path(DATA_DIR, "gts_ols_reduction_de.csv"))
```

---

## Spain (ES)

```{r es-subset, echo=FALSE}
target_country <- "ES"

sub_df <- daily_clean %>%
  filter(country_code == target_country) %>%
  dplyr::select(
    all_of(TARGET_LOAD_COL), temperature_c, dewpoint_c, wind_speed_ms,
    solar_radiation_wm2, tcc, precipitation_mm, month, is_weekend
  ) %>%
  na.omit()
```

### Model 1: Full General Uncentered OLS Baseline

```{r es-model1-baseline, echo=FALSE}
protected_vars <- c("temperature_c", "I(temperature_c^2)", "month", "is_weekend")
candidate_vars_full <- c("dewpoint_c", "wind_speed_ms", "solar_radiation_wm2", "tcc", "precipitation_mm")
current_vars_m1 <- c(protected_vars, candidate_vars_full)
formula_m1 <- paste(TARGET_LOAD_COL, "~", paste(current_vars_m1, collapse = " + "))
model_ols_general <- lm(as.formula(formula_m1), data = sub_df)
vif(model_ols_general) # Check for multicollinearity
summary(model_ols_general)
```

#### Omit `solar_radiation_wm2` due to high collinearity

```{r es-311a}
linearHypothesis(model_ols_general, c("solar_radiation_wm2 = 0"))
```

```{r es-general-reduced-solar_radiation_wm2}
model_ols_reduced1 <- update(model_ols_general, . ~ . - solar_radiation_wm2)
summary(model_ols_reduced1)
```

#### Jointly significant test of all insignificant variables

```{r es-311b}
linearHypothesis(model_ols_general, c("solar_radiation_wm2 = 0", "month02 = 0", "month07 = 0", "dewpoint_c = 0"))
```

#### Omit `month02`, `month07`, `dewpoint_c` due to insignificance

```{r es-month11-reduced}
sub_df_reduced <- sub_df %>%
  mutate(
    month_reduced = fct_collapse(as.factor(month), "01" = c("02", "07"))
  )

model_ols_reduced2 <- update(
  model_ols_general, 
  . ~ . - dewpoint_c - solar_radiation_wm2 - month + month_reduced,
  data = sub_df_reduced
)

summary(model_ols_reduced2)
```

**Final model for Spain:** `load_mw ~ temperature_c + I(temperature_c^2) + is_weekend + wind_speed_ms + tcc + month_reduced + precipitation_mm` (February and July merged into January baseline).

### OLS Diagnostic Summary Table

```{r es-ols-summary, echo=FALSE}
coefs_m2 <- coef(model_ols_reduced2)
r_sq_m2 <- summary(model_ols_reduced2)$r.squared
sub_df_reduced$model_ols_reduced2 <- residuals(model_ols_reduced2)

summary_list_m2 <- data.frame(
  Country = target_country,
  Beta_Intercept = coefs_m2[1],
  Beta_T = coefs_m2[2],
  Beta_T2 = coefs_m2[3],
  R_Squared = r_sq_m2,
  row.names = NULL
)
knitr::kable(summary_list_m2, caption = "Reduced OLS Framework Summary Parameters")
```

### Model 2: Centered Specific OLS Framework

```{r es-ols-centered, echo=FALSE}
sub_df_reduced$temperature_c_centered <- scale(sub_df_reduced$temperature_c, scale = FALSE)
model_ols_centered <- update(
  model_ols_reduced2, 
  . ~ . - temperature_c - I(temperature_c^2) + temperature_c_centered + I(temperature_c_centered^2),
  data = sub_df_reduced
)
cat("Final specification VIF\n")
vif(model_ols_centered)
summary(model_ols_centered)
```

### AIC Model Comparison

```{r es-AIC, echo=FALSE}
aic_output <- AIC(model_ols_centered, model_ols_reduced2)
knitr::kable(aic_output, caption = "Model Fitness Selection Matrix (Lower AIC wins)")
```

---

## Finland (FI)

```{r fi-subset, echo=FALSE}
target_country <- "FI"

sub_df <- daily_clean %>%
  filter(country_code == target_country) %>%
  dplyr::select(
    all_of(TARGET_LOAD_COL), temperature_c, dewpoint_c, wind_speed_ms,
    solar_radiation_wm2, tcc, precipitation_mm, month, is_weekend
  ) %>%
  na.omit()
```

### Model 1: Full General Uncentered OLS Baseline

```{r fi-model1-baseline, echo=FALSE}
protected_vars <- c("temperature_c", "I(temperature_c^2)", "month", "is_weekend")
candidate_vars_full <- c("dewpoint_c", "wind_speed_ms", "solar_radiation_wm2", "tcc", "precipitation_mm")
current_vars_m1 <- c(protected_vars, candidate_vars_full)
formula_m1 <- paste(TARGET_LOAD_COL, "~", paste(current_vars_m1, collapse = " + "))
model_ols_general <- lm(as.formula(formula_m1), data = sub_df)
vif(model_ols_general) # Check for multicollinearity
summary(model_ols_general)
```

#### Omit `dewpoint_c`, `solar_radiation_wm2` due to high collinearity

```{r fi-411a}
linearHypothesis(model_ols_general, c("dewpoint_c = 0", "solar_radiation_wm2 = 0"))
```

```{r fi-general-reduced-dewpoint_c-solar_radiation_wm2}
model_ols_reduced1 <- update(model_ols_general, . ~ . - dewpoint_c - solar_radiation_wm2)
summary(model_ols_reduced1)
```

#### Jointly significant test of insignificant variables

```{r fi-412a}
linearHypothesis(model_ols_general, c("dewpoint_c = 0", "solar_radiation_wm2 = 0", "month12 = 0", "tcc = 0", "precipitation_mm = 0"))
```

```{r fi-general-reduced2}
sub_df_reduced <- sub_df %>%
  mutate(
    month_reduced = fct_collapse(as.factor(month), "01" = c("12"))
  )

model_ols_reduced2 <- update(
  model_ols_general, 
  . ~ . - dewpoint_c - solar_radiation_wm2 - month + month_reduced - precipitation_mm - tcc,
  data = sub_df_reduced
)

summary(model_ols_reduced2)
```

**Final model for Finland:** `load_mw ~ temperature_c + I(temperature_c^2) + is_weekend + wind_speed_ms + month_reduced` (December merged into January baseline).

### OLS Diagnostic Summary Table

```{r fi-ols-summary, echo=FALSE}
coefs_m2 <- coef(model_ols_reduced2)
r_sq_m2 <- summary(model_ols_reduced2)$r.squared
sub_df_reduced$model_ols_reduced2 <- residuals(model_ols_reduced2)

summary_list_m2 <- data.frame(
  Country = target_country,
  Beta_Intercept = coefs_m2[1],
  Beta_T = coefs_m2[2],
  Beta_T2 = coefs_m2[3],
  R_Squared = r_sq_m2,
  row.names = NULL
)
knitr::kable(summary_list_m2, caption = "Reduced OLS Framework Summary Parameters")
```

### Model 2: Centered Specific OLS Framework

```{r fi-ols-centered, echo=FALSE}
sub_df_reduced$temperature_c_centered <- scale(sub_df_reduced$temperature_c, scale = FALSE)
model_ols_centered <- update(
  model_ols_reduced2, 
  . ~ . - temperature_c - I(temperature_c^2) + temperature_c_centered + I(temperature_c_centered^2),
  data = sub_df_reduced
)
cat("Final specification VIF\n")
vif(model_ols_centered)
summary(model_ols_centered)
```

### AIC Model Comparison

```{r fi-AIC, echo=FALSE}
aic_output <- AIC(model_ols_centered, model_ols_reduced2)
knitr::kable(aic_output, caption = "Model Fitness Selection Matrix (Lower AIC wins)")
```

---

## France (FR)

```{r fr-subset, echo=FALSE}
target_country <- "FR"

sub_df <- daily_clean %>%
  filter(country_code == target_country) %>%
  dplyr::select(
    all_of(TARGET_LOAD_COL), temperature_c, dewpoint_c, wind_speed_ms,
    solar_radiation_wm2, tcc, precipitation_mm, month, is_weekend
  ) %>%
  na.omit()
```

### Model 1: Full General Uncentered OLS Baseline

```{r fr-model1-baseline}
protected_vars <- c("temperature_c", "I(temperature_c^2)", "month", "is_weekend")
candidate_vars_full <- c("dewpoint_c", "wind_speed_ms", "solar_radiation_wm2", "tcc", "precipitation_mm")
current_vars_m1 <- c(protected_vars, candidate_vars_full)
formula_m1 <- paste(TARGET_LOAD_COL, "~", paste(current_vars_m1, collapse = " + "))
model_ols_general <- lm(as.formula(formula_m1), data = sub_df)
vif(model_ols_general) # Check for multicollinearity
summary(model_ols_general)
```


#### Jointly significant test of insignificant variables

```{r fr-411}
linearHypothesis(model_ols_general, c("tcc = 0", "precipitation_mm = 0", "wind_speed_ms = 0"))
```

```{r fr-reduced1}
model_ols_reduced1 <- update(model_ols_general, . ~ . - tcc - precipitation_mm - wind_speed_ms)
summary(model_ols_reduced1)
```

**Final model for France:** `load_mw ~ temperature_c + I(temperature_c^2) + is_weekend + dewpoint_c + solar_radiation_wm2 + month`

### OLS Diagnostic Summary Table

```{r fr-ols-summary, echo=FALSE}
coefs_m2 <- coef(model_ols_reduced1)
r_sq_m2 <- summary(model_ols_reduced1)$r.squared
sub_df$model_ols_reduced1 <- residuals(model_ols_reduced1)

summary_list_m2 <- data.frame(
  Country = target_country,
  Beta_Intercept = coefs_m2[1],
  Beta_T = coefs_m2[2],
  Beta_T2 = coefs_m2[3],
  R_Squared = r_sq_m2,
  row.names = NULL
)
knitr::kable(summary_list_m2, caption = "Reduced OLS Framework Summary Parameters")
```

### Model 2: Centered Specific OLS Framework

```{r fr-ols-centered, echo=FALSE}
sub_df$temperature_c_centered <- scale(sub_df$temperature_c, scale = FALSE)
model_ols_centered <- update(
  model_ols_reduced1, 
  . ~ . - temperature_c - I(temperature_c^2) + temperature_c_centered + I(temperature_c_centered^2),
  data = sub_df
)
cat("Final specification VIF\n")
vif(model_ols_centered)
summary(model_ols_centered)
```

### AIC Model Comparison

```{r fr-AIC, echo=FALSE}
aic_output <- AIC(model_ols_centered, model_ols_reduced2)
knitr::kable(aic_output, caption = "Model Fitness Selection Matrix (Lower AIC wins)")
```


---

## Poland (PL)

```{r pl-subset, echo=FALSE}
target_country <- "PL"

sub_df <- daily_clean %>%
  filter(country_code == target_country) %>%
  dplyr::select(
    all_of(TARGET_LOAD_COL), temperature_c, dewpoint_c, wind_speed_ms,
    solar_radiation_wm2, tcc, precipitation_mm, month, is_weekend
  ) %>%
  na.omit()
```

### Model 1: Full General Uncentered OLS Baseline

```{r pl-model1-baseline, echo=FALSE}
protected_vars <- c("temperature_c", "I(temperature_c^2)", "month", "is_weekend")
candidate_vars_full <- c("dewpoint_c", "wind_speed_ms", "solar_radiation_wm2", "tcc", "precipitation_mm")
current_vars_m1 <- c(protected_vars, candidate_vars_full)
formula_m1 <- paste(TARGET_LOAD_COL, "~", paste(current_vars_m1, collapse = " + "))
model_ols_general <- lm(as.formula(formula_m1), data = sub_df)
vif(model_ols_general) # Check for multicollinearity
summary(model_ols_general)
```

#### Jointly significant test

```{r pl-611}
linearHypothesis(model_ols_general, c("month03 = 0", "month10 = 0", "month11 = 0", "wind_speed_ms = 0", "solar_radiation_wm2 = 0", "tcc = 0", "precipitation_mm = 0"))
```

```{r pl-general-reduced}
sub_df_reduced <- sub_df %>%
  mutate(
    month_reduced = fct_collapse(as.factor(month), "01" = c("03", "10", "11"))
  )

model_ols_reduced1 <- update(
  model_ols_general, 
  . ~ . - wind_speed_ms - solar_radiation_wm2 - tcc - precipitation_mm - month + month_reduced,
  data = sub_df_reduced
)

summary(model_ols_reduced1)
```

**Final model for Poland:** `load_mw ~ temperature_c + I(temperature_c^2) + is_weekend + dewpoint_c + month_reduced` (March, October, November merged into January baseline).

### OLS Diagnostic Summary Table

```{r pl-ols-summary, echo=FALSE}
coefs_m2 <- coef(model_ols_reduced1)
r_sq_m2 <- summary(model_ols_reduced1)$r.squared
sub_df_reduced$model_ols_reduced1 <- residuals(model_ols_reduced1)

summary_list_m2 <- data.frame(
  Country = target_country,
  Beta_Intercept = coefs_m2[1],
  Beta_T = coefs_m2[2],
  Beta_T2 = coefs_m2[3],
  R_Squared = r_sq_m2,
  row.names = NULL
)
knitr::kable(summary_list_m2, caption = "Reduced OLS Framework Summary Parameters")
```

### Model 2: Centered Specific OLS Framework

```{r pl-ols-centered, echo=FALSE}
sub_df_reduced$temperature_c_centered <- scale(sub_df_reduced$temperature_c, scale = FALSE)
model_ols_centered <- update(
  model_ols_reduced1, 
  . ~ . - temperature_c - I(temperature_c^2) + temperature_c_centered + I(temperature_c_centered^2),
  data = sub_df_reduced
)
cat("Final specification VIF\n")
vif(model_ols_centered)
summary(model_ols_centered)
```

### AIC Model Comparison

```{r pl-AIC, echo=FALSE}
aic_output <- AIC(model_ols_centered, model_ols_reduced1)
knitr::kable(aic_output, caption = "Model Fitness Selection Matrix (Lower AIC wins)")
```

---

## Variable Selection Summary

| Country | Retained Weather Regressors | Dropped (collinear / insignificant) | Month Merges |
|---------|----------------------------|--------------------------------------|--------------|
| DE | `wind_speed_ms`, `tcc` | `dewpoint_c`, `solar_radiation_wm2`, `precipitation_mm` | Nov → Jan |
| ES | `wind_speed_ms`, `tcc`, `precipitation_mm` | `solar_radiation_wm2`, `dewpoint_c` | Feb, Jul → Jan |
| FI | `wind_speed_ms` | `dewpoint_c`, `solar_radiation_wm2`, `precipitation_mm` | Dec → Jan |
| FR | `dewpoint_c`, `solar_radiation_wm2` | `wind_speed_ms`, `tcc`, `precipitation_mm` | None |
| PL | `dewpoint_c` | `wind_speed_ms`, `solar_radiation_wm2`, `tcc`, `precipitation_mm` | Mar, Oct, Nov → Jan |

```{r gts-csv-export, echo=FALSE}
write_csv(GTS_OLS_SURVIVING, file.path(DATA_DIR, "gts_ols_surviving.csv"))
write_csv(GTS_MONTH_MERGES, file.path(DATA_DIR, "gts_month_merges.csv"))
```

# Main Model: ARDL

> **Purpose.** Fix the autocorrelation found in section 4 by stepping through three nested specifications: a Distributed-Lag (DL) of weather only, then an ARDL adding lags of the dependent variable, then a **seasonal ARDL** that adds the lag-7 terms motivated by the weekly spike in the OLS residual ACF.
>
> **Expected result.** Each step reduces the Breusch-Godfrey p-values at orders 1 **and** 7. The seasonal-ARDL with lag 7 on both `d(log_load)` and `d(temperature_c)` should have the lowest AIC and a residual ACF without visible spikes at any lag. Sign of the short-run temperature effect (sum of `d(temp)` coefs) should match climate: negative in heating-dominated countries (`FI`), positive in cooling-dominated countries (`ES`).
>
> **If not as expected.** If BG at order 7 still rejects after the seasonal specification, add `L(d(log_load), 14)` (two-week lag) or use HAC-robust standard errors via `coeftest(model, vcov = vcovHAC(model))`.

Following standard ARDL practice, we work on **daily** data in **differences** of `log(load)` and `temperature_c`. Weekday fixed effects are encoded as one-hot dummies inside the `zoo` object.

```{r ardl-focus, results='asis', echo=FALSE}
ardl_focus_results <- list()

for (cc in FOCUS_COUNTRIES) {
  cat("\n\n## ARDL -", cc, "\n\n")
  sub <- daily_clean %>% filter(country_code == cc)
  z <- to_daily_zoo(sub)
  mods <- fit_ardl_models(z, country_code = cc)

  cat("### Step 1 - Distributed Lag (weather only)\n")
  cat("BG check on DL parsimonious: ")
  print(bgtest(residuals(mods$dl_pars) ~ 1, order = 7))

  cat("\n### Step 2 - ARDL (add lags of dependent)\n")
  cat("ANOVA full vs parsimonious ARDL (H0: dropped lags are jointly zero):\n")
  print(anova(mods$ardl_full, mods$ardl_pars))

  cat("\n### Step 3 - Seasonal ARDL (lag 7 on dep and temperature)\n")
  cat("ANOVA cannot be used between ardl_pars and ardl_seasonal because the lag-7 terms drop 5 extra initial observations; compare via AIC + BG at order 7 instead.\n")
  bg7_pars <- bgtest(residuals(mods$ardl_pars) ~ 1, order = 7)$p.value
  bg7_seas <- bgtest(residuals(mods$ardl_seasonal) ~ 1, order = 7)$p.value

  aic_tab <- data.frame(
    model = c("dl_pars", "ardl_pars", "ardl_seasonal"),
    AIC = c(AIC(mods$dl_pars), AIC(mods$ardl_pars), AIC(mods$ardl_seasonal)),
    bg_p_order7 = c(
      bgtest(residuals(mods$dl_pars) ~ 1, order = 7)$p.value,
      bg7_pars,
      bg7_seas
    )
  )
  cat("\n**AIC and weekly-autocorrelation comparison**\n\n")
  print(knitr::kable(aic_tab, digits = c(0, 2, 4)))

  model <- mods$ardl_seasonal
  cat("\n### Preferred (seasonal ARDL) coefficients\n")
  print(summary(model))

  cat("\n### HAC-robust coefficients (Newey-West)\n")
  print(coeftest(model, vcov = vcovHAC(model)))

  cat("\n### Breusch-Godfrey p-values\n")
  for (k in c(1, 5, 7, 14)) {
    p <- bgtest(residuals(model) ~ 1, order = k)$p.value
    cat(sprintf("BG order %2d: p = %.4f\n", k, p))
  }

  acf(residuals(model), main = paste("Seasonal ARDL residuals ACF -", cc))

  diag <- ardl_diagnostics(model)
  ardl_focus_results[[cc]] <- tibble(
    country = cc,
    short_run_temp_sum = diag$short_run_temp,
    bg_p_order1 = diag$bg[["bg_p_order1"]],
    bg_p_order7 = diag$bg[["bg_p_order7"]],
    bg_p_order14 = diag$bg[["bg_p_order14"]],
    aic = diag$aic
  )
}

ardl_focus_table <- bind_rows(ardl_focus_results)
ardl_focus_table %>%
  knitr::kable(digits = 4, caption = "Seasonal ARDL summary - focus countries")
```

**Short-run temperature effect:** Sum of coefficients on `d(temperature_c)` and its lags (contemporaneous and lag 1, 3, 7). Excludes `d(temp_c2)` because that captures only the nonlinearity, not the marginal degree-Celsius response.

# ARDL Diagnostics, General-to-Specific Table, and ARDL vs ARIMA (DE)

Course checklist (ARDL section): White test, RESET on final ARDL, general-to-specific table, and comparison of final ARDL with ARIMA on the same daily sample.

```{r ardl-checklist-diagnostics, echo=FALSE}
ardl_diag_focus <- map_dfr(FOCUS_COUNTRIES, function(cc) {
  sub <- daily_clean %>% filter(country_code == cc)
  z <- to_daily_zoo(sub)
  mods <- fit_ardl_models(z, country_code = cc)
  m <- mods$ardl_seasonal
  bind_cols(
    tibble(country = cc, short_run_temp = short_run_temp_effect(m)),
    as_tibble(as.list(ardl_final_diagnostics(m)))
  )
})

write_csv(ardl_diag_focus, file.path(DATA_DIR, "ardl_diagnostics_focus.csv"))

ardl_diag_focus %>%
  select(country, short_run_temp, bp_p, white_p, reset_p, bg_p_order1, bg_p_order7, aic, bic) %>%
  knitr::kable(digits = 4, caption = "Final seasonal ARDL diagnostics - focus countries (BP, White, RESET, BG)")
```

```{r ardl-gts-de, echo=FALSE}
# General-to-specific model selection table for Germany (flagship country).
de_sub <- daily_clean %>% filter(country_code == "DE")
z_de <- to_daily_zoo(de_sub)
mods_de <- fit_ardl_models(z_de, country_code = "DE")

gts_de <- bind_rows(
  model_step_summary(mods_de$dl_full, "DL full"),
  model_step_summary(mods_de$dl_pars, "DL parsimonious"),
  model_step_summary(mods_de$ardl_full, "ARDL full"),
  model_step_summary(mods_de$ardl_pars, "ARDL parsimonious"),
  model_step_summary(mods_de$ardl_seasonal, "Seasonal ARDL (final)")
)

write_csv(gts_de, file.path(DATA_DIR, "ardl_model_comparison_de.csv"))

gts_de %>%
  knitr::kable(digits = 4, caption = "General-to-specific ARDL selection - Germany")
```

```{r ardl-vs-arima-daily-de, echo=FALSE}
# Compare final seasonal ARDL with ARIMA on the same daily Δlog(load) for DE.
de_z <- to_daily_zoo(de_sub)
dlog <- diff(de_z$log_load)
dtemp <- diff(de_z$temperature_c)
dtemp2 <- diff(de_z$temp_c2)
wd_cols <- grep("^wd_", colnames(de_z), value = TRUE)
n_z <- nrow(de_z)
wd_for_diff <- as.matrix(de_z[2:n_z, wd_cols, drop = FALSE])
# Align all series after differencing (dynlm drops first obs; lag-7 drops more).
align_start <- 8L
y_arima <- as.numeric(dlog)[align_start:length(dlog)]
xreg_mat <- cbind(
  dtemp = as.numeric(dtemp)[align_start:length(dtemp)],
  dtemp2 = as.numeric(dtemp2)[align_start:length(dtemp2)],
  wd_for_diff[align_start:nrow(wd_for_diff), , drop = FALSE]
)

fit_arima_daily <- auto.arima(
  y_arima,
  xreg = xreg_mat,
  seasonal = FALSE,
  stepwise = TRUE,
  approximation = TRUE,
  max.p = 3,
  max.q = 3,
  max.order = 8
)

ardl_final_de <- mods_de$ardl_seasonal
ardl_vs_arima <- tibble(
  model = c("Seasonal ARDL (final)", "ARIMA + weather xreg"),
  aic = c(AIC(ardl_final_de), AIC(fit_arima_daily)),
  bic = c(BIC(ardl_final_de), BIC(fit_arima_daily)),
  arima_order = c(
    NA_character_,
    paste0("(", paste(fit_arima_daily$arma[c(1, 6, 2)], collapse = ","), ")")
  )
)

write_csv(ardl_vs_arima, file.path(DATA_DIR, "ardl_vs_arima_de.csv"))

cat("Selected daily ARIMA for DE:\n")
print(fit_arima_daily)
ardl_vs_arima %>%
  knitr::kable(digits = 2, caption = "Final ARDL vs ARIMA - daily Germany (same Δlog load sample)")
```

**Long-run interpretation:** Because daily log load is I(1), the levels relationship between load and temperature is captured primarily through the static OLS U-curve (Section 4). The ARDL in differences identifies short-run responses to weather shocks; a full long-run multiplier would require a levels ARDL or an error-correction specification when variables co-integrate. Given mixed integration properties of temperature across countries, we report short-run sums in the table above and discuss long-run nonlinearity via OLS turning points.

# Robustness: SARIMAX / ARIMA with Weather (Germany, Hourly)

> **Purpose.** Out-of-sample forecast test using ARIMA on **hourly** DE log load with temperature and hour-of-day dummies as exogenous regressors.
>
> **Expected result.** `auto.arima` + xreg beats the seasonal-naive 24-hour benchmark on MAE, RMSE and MAPE (computed on MW levels). The fitted ARIMA picks up at least one autoregressive and one moving-average term and a seasonal differencing at lag 24.
>
> **If not as expected.** If the naive benchmark wins, replace the hour dummies with Fourier terms `forecast::fourier(K = 2, ts(..., frequency = 24*7))` to model weekly seasonality, or force `Arima(..., seasonal = list(order = c(0, 1, 1), period = 24))` explicitly.
>
> **What to look at.** Forecast metrics table (MAE / RMSE / MAPE), the holdout-period forecast vs actual plot, and the AIC of the chosen ARIMA.

```{r sarimax-de, cache=TRUE, echo=FALSE}
de_hourly <- hourly_raw %>%
  mutate(
    timestamp = as.POSIXct(timestamp, tz = "UTC"),
    date = as.Date(timestamp),
    country_code = as.character(country_code),
    hour = as.integer(format(timestamp, "%H"))
  ) %>%
  filter(
    country_code == "DE",
    date <= ANALYSIS_END,
    is.finite(load_mw),
    is.finite(temperature_c)
  ) %>%
  arrange(timestamp) %>%
  mutate(
    log_load = log(load_mw),
    hour_fe = factor(hour)
  )

# Design matrix: temperature + hour-of-day dummies
xreg_full <- model.matrix(~ temperature_c + hour_fe, data = de_hourly)[, -1, drop = FALSE]
y <- de_hourly$log_load

h_holdout <- 14 * 24
n <- length(y)
y_train <- y[1:(n - h_holdout)]
xreg_train <- xreg_full[1:(n - h_holdout), , drop = FALSE]
y_test <- y[(n - h_holdout + 1):n]
xreg_test <- xreg_full[(n - h_holdout + 1):n, , drop = FALSE]

# Seasonal ARIMA with exogenous regressors (daily seasonality period = 24).
fit_arima_xreg <- auto.arima(
  ts(y_train, frequency = 24),
  xreg = xreg_train,
  seasonal = TRUE,
  stepwise = TRUE,
  approximation = TRUE,
  max.p = 2, max.q = 2,  # Limit non-seasonal orders
  max.P = 1, max.Q = 1   # Limit seasonal orders strictly
)

cat("Selected model:\n")
print(fit_arima_xreg)

fc <- forecast(fit_arima_xreg, h = h_holdout, xreg = xreg_test)
fc_mean <- as.numeric(fc$mean)

# Seasonal-naive benchmark at period 24 (yesterday's hour-of-day pattern).
naive_fc <- as.numeric(
  forecast::snaive(ts(y_train, frequency = 24), h = h_holdout)$mean
)

# Metrics are reported on MW (level) units so MAPE has a physical meaning.
forecast_metrics <- function(actual_log, pred_log) {
  actual <- exp(actual_log)
  pred <- exp(pred_log)
  tibble(
    MAE_MW = mean(abs(actual - pred)),
    RMSE_MW = sqrt(mean((actual - pred)^2)),
    MAPE = mean(abs((actual - pred) / actual))
  )
}

fc_eval <- bind_rows(
  forecast_metrics(y_test, fc_mean) %>% mutate(model = "ARIMA + weather xreg"),
  forecast_metrics(y_test, naive_fc) %>% mutate(model = "Seasonal naive (period 24)")
) %>%
  select(model, everything())

fc_eval %>%
  knitr::kable(
    digits = 4,
    caption = "Out-of-sample forecast accuracy - DE hourly load (metrics on MW levels)"
  )

write_csv(fc_eval, file.path(DATA_DIR, "forecast_metrics_de.csv"))
```

```{r sarimax-plot, fig.height=5, echo=FALSE}
plot_idx <- (n - h_holdout - 7 * 24):n
plot(
  plot_idx, y[plot_idx],
  type = "l", col = "gray40",
  main = "DE hourly log(load): holdout forecast (last 14 days)",
  xlab = "Observation index", ylab = "log(load)"
)
holdout_idx <- (n - h_holdout + 1):n
lines(holdout_idx, y_test, col = "black", lwd = 1.5)
lines(holdout_idx, fc_mean, col = "red", lwd = 1.5)
legend(
  "topleft",
  legend = c("Actual (holdout)", "ARIMA + xreg forecast"),
  col = c("black", "red"),
  lty = 1,
  bty = "n"
)
```

```{r export-report-figures, echo=FALSE}
source("R/export-report-figures.R")
holdout_idx <- (n - h_holdout + 1):n
plot_idx <- (n - h_holdout - 7 * 24):n
export_report_figures(
  daily_clean = daily_clean,
  hourly_raw = hourly_raw,
  focus_countries = FOCUS_COUNTRIES,
  analysis_end = ANALYSIS_END,
  fig_dir = FIG_DIR,
  y_test = y_test,
  fc_mean = fc_mean,
  y_sub_for_plot = y,
  holdout_idx = holdout_idx,
  plot_idx = plot_idx
)
```

# Appendix: Cross-Country Summary Table

> **Purpose.** Confirm that the focus-country findings generalise across the 21-country panel - one row per country with the U-curve turning point, short-run temperature effect from the seasonal ARDL, and BG residual diagnostic.
>
> **Expected result.** Negative short-run temperature effect in cold-climate countries (`SE`, `FI`, `PL`), positive in summer-cooling countries (`ES`, `GR`, `IT`, `PT`). Most countries should show `bg_p_order1 < 0.10` even after the seasonal ARDL.
>
> **If not as expected.** Countries with extreme outlier coefficients usually have data quality issues - partial ENTSO-E coverage, weather centroid placed offshore, or load gaps. Note these in the report rather than re-running the pipeline.

```{r appendix-all-countries, echo=FALSE}
appendix_metrics <- map_dfr(AVAILABLE_COUNTRIES, function(cc) {
  sub <- daily_clean %>% filter(country_code == cc)
  if (nrow(sub) < 100) return(NULL)

  m_ols <- fit_baseline_ols(sub)
  tp <- u_curve_turning_point(m_ols)

  z <- to_daily_zoo(sub)
  mods <- tryCatch(fit_ardl_models(z, country_code = cc), error = function(e) NULL)
  if (is.null(mods)) {
    sr_temp <- NA_real_
    ardl_aic <- NA_real_
    bg1 <- NA_real_
  } else {
    m <- mods$ardl_seasonal
    d <- ardl_diagnostics(m)
    sr_temp <- d$short_run_temp
    ardl_aic <- d$aic
    bg1 <- d$bg[["bg_p_order1"]]
  }

  tibble(
    country = cc,
    n_days = nrow(sub),
    mean_temp_c = mean(sub$temperature_c),
    ols_r2 = summary(m_ols)$r.squared,
    u_curve_turning_point_c = tp,
    ardl_short_run_temp = sr_temp,
    ardl_bg_p_order1 = bg1,
    ardl_aic = ardl_aic
  )
})

write_csv(appendix_metrics, file.path(DATA_DIR, "appendix_country_summary.csv"))

appendix_metrics %>%
  arrange(country) %>%
  knitr::kable(digits = 4, caption = "All-country summary (appendix)")
```

# Session Information

> **Purpose.** Capture R version and package versions so the analysis is reproducible.
>
> **Expected result.** R 4.5 or newer with all packages from the `required_pkgs` list reported.

```{r session-info}
sessionInfo()
```
